---
title: "General Social Survey"
author: "Diego Martinez"
date: "7/5/22"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```




#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  
First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

$H_0$: There is no relationship between the opinion of death penalty & gun ownership. They are independent. 

$H_A$: There is a relationship between the opinion of death penalty & gun ownership. They are dependent. 

## Methods

They're both categorical variables with 2 levels. Thus, we will use a chi-squared test of independence and a fisher exact test for odds ratio. OWNGUN has YES/NO. CAPPUN has YES/NO. 

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

## Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results


```{r}
barchartGC(~CAPPUN + OWNGUN,data=GSSdata)
barchartGC(~CAPPUN + OWNGUN,data=GSSdata, type="percent")

```

There is a clear difference. The pink is in favor by 60. While the blue is in favor by 55. The pink is opposed by 45. While the blue is opposed by about 20. The pink is in favor more by 5. While the blue is more opposed by 25. 


<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

```{r}
table1 <- xtabs(~CAPPUN + OWNGUN, data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

This table shows that if you believe in capital punishment you are likely not in favor of either yes or no. If you are oppose of capital punishment you are likely to say no 71% of the time. This table shows that if you believe in owning a gun  you are likely to say yes 18% more of the time. If you are against owning a gun you are 18% more likely to say no. 

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for table1 is  0.02022. The chi square statistic is 5.3932 with a degree of freedom of 1. The chi-square statistic p-value is .01651 which is the same p-value as the fisher exact test. The odds ratio is 2.27 to 1.  Therefore the odds are 2.27 to 1 likelier that somebody who has a gun will be in favor of capital punishment. We can reject the null hypothesis because the p-value is too small.  


<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

$H_0$: There is no difference between years of education between those who did or did not vote in 2012. 

$H_A$: There is a difference between years of education between those who did or did not vote in 2012. 


## Methods


This is a Cat vs Num. Since the difference in years is numerical & if they did or did not vote is categorical. We will use a box plot, favstats, and a t.test for this question. 


<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

## Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

```{r}
plot(EDUC~VOTE16,data=GSSdata)
```

The median for the voted category is 15. The median for the did not vote category is 13. There is 3 outliers in the did not vote category. The 1.5 IQR for the voted box plot is a lot lengthier than the did not vote. The IQR for the Voted category is much higher than the did not voted. Currently, there is clearly a difference between the two. 


<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results


```{r}
favstats(EDUC~VOTE16,data=GSSdata)
```

The voted category has a mean of 15.625 and a median of 15. The did not vote category has a mean of 13.68 and a median is 13. This clearly states that there is a difference between the two even more. The n value for the voted category is 125. The did not vote category is 50. 

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

```{r}
t.test(EDUC~VOTE16,data=GSSdata)
```

The t-value is 4.598. The degrees of freedom is 102.27. The p-value is 0.00001224. Thus, we can reject the null hypothesis since the p-value is too small. There is clearly a difference between the two variables. 

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  

$H_A$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE 

## Methods

We are explaining that there is a significant difference in average score on WORDSUM tests between at least two levels of degree. We will be examining a categorical variable cultivation. We will also use a T-test to compare the average score on the WORDSUM test between at least two levels of degree. The P-value will determine if whether or not we're correct: it has to be lower than .05. The WORDSUM is numerical. Degree is categorical.

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

## Results  

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

```{r}
boxplot(WORDSUM~DEGREE,data=GSSdata)
```

The median for the bachelor degree is a bit higher than the other two medians. 
The bachelor degree median is about 10, high school degree median is about 9, and the LT high school median is about 8. The bachelor degree box plot doesn't have any outliers, highschool degree has 2 outliers below it's minimum, LT highschool degree has 1 outlier below the minimum & one above the max. So far based off of this box plot, the bachelor degree (categorical) has a higher median for the WORDSUM score (numerical) out of all 3 degrees. 

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

```{r}
favstats(WORDSUM~DEGREE,data=GSSdata)
```

The median for a bachelor degree is 10, 11 for a graduate, 9 for a high school degree, 10 for a junior college, and a 7 for LT high school. The max is the same for 4 out of the 5 degrees (11 for LT). The mean is slightly different for all of them. Overall, the graduate has a higher median and mean. While having an equal maximum. It's the standout and has the highest average out of all the degrees. LT high school appears to have the lowest median, maximum, and average. 

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSSdata)
summary.aov(model1)
```

We can infer that the P-value is .000236. From the AOV, we can reject the null hypothesis due to how small the p-value is. If the p-value is not .05 or higher, we can reject it due to how small the chances are. In every problem, we know we're accurate when we can reject the null hypothesis. There is a difference between degrees and how they score (on average) in the WORDSUM. The graduates clearly pull off a greater average than any of the 5 degrees provided with data. 

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion


Question 1 ended up with a p-value of  0.0202 so we reject null hypothesis for this question. There is a difference between these two categorical variables. Question 2 ended up with a p-value of 0.00001224 so we reject the null hypothesis for this question. There is a difference between the numerical and categorical variable. Question 3 ended up with a p-value of .000236 so we reject the null hypothesis. There is a difference between the numerical and categorical variable. In conclusion, I was able to reject the null hypothesis for all 3 questions & able to fully complete each step towards rejecting the null hypothesis. 


<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

